Layered Formal Verification of a TCP Stack

### 2021-08-30 SecDev21/paper.tex
clarifications

Using KLEE on the function that process incoming segments helped us to make faster the verification of the user function. Applied to the user function, this method would not have been enough. KLEE can only unroll loops when the loops present in the user functions can finish after a high number of iterations which may leads to a combinatorial explosion in the number of paths explored by KLEE. KLEE has other limitations that were useful in SPARK for our verification like the detection of memory leaks.

@@

\subsubsection{Memory leak}

@@

In total, for this work, 50 functions have been translated from C to SPARK. This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove the TCP specification. All the socket functions and the user functions of TCP have been translated and proved in SPARK, in particular \spark{TCP_Init}, \spark{TCP_Connect}, \spark{TCP_Listen}, \spark{TCP_Send}, \spark{TCP_Receive}, \spark{TCP_Shutdown}, \spark{TCP_Accept} and \spark{TCP_Abort}. The corresponding socket function that makes the interface between Socket function have also been fully translated to SPARK.
At the end, 25\% of the functions that implement the TCP protocol have been rewritten in SPARK.

>>>>

Our choice of combining deductive verification with SPARK and symbolic execution with KLEE was driven by practical considerations which are typical of industrial applications. Using KLEE on user functions would be difficult due to the presence of loops with a possible high number of iterations, which symbolic execution will need to unroll, leading to combinatorial explosion. Using SPARK on the function that processes incoming segments would be desirable, and may be done in the future, but was not doable in the timeframe of this project. Note that KLEE has other limitations compared to SPARK in terms of the verification scope. For example it is lacking a detection of memory leaks, which turned out to be useful in SPARK, see section~\ref{sec:memory-leak}.

@@

\label{sec:memory-leak}
\subsubsection{Memory leak}

@@

In total, for this work, 50 functions have been translated from C to SPARK. This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove the TCP specification. All the socket functions and the user functions of TCP have been translated and proved in SPARK, in particular \spark{TCP_Init}, \spark{TCP_Connect}, \spark{TCP_Listen}, \spark{TCP_Send}, \spark{TCP_Receive}, \spark{TCP_Shutdown}, \spark{TCP_Accept} and \spark{TCP_Abort}. The corresponding socket functions from the Socket interface have also been fully translated to SPARK.
At the end, 25\% of the functions that implement the TCP protocol have been rewritten in SPARK.
